\documentclass[fleqn,11pt]{article}
\usepackage{sty/fullpage}

% to recognize url address 
\usepackage{url}

% to order references automatically
\usepackage{sty/cite}

% \usepackage{sty/psfig}
% \usepackage{subfigure}
%\usepackage{lib/tex/defpaper2e}

\newcommand{\qed}{\hfill \rule{2mm}{3mm}}
\newcommand{\eat}[1]{}
\newcommand{\lbok}{\linebreak[0]}
\newcommand{\reminder}[1]{ [[[ \marginpar{\mbox{$<==$}} #1 ]]] }
\newcommand{\eatreminders}{\renewcommand{\reminder}[1]{}}
\newcommand{\comment}[1]{ {\footnotesize {#1} } }
\newcommand{\QED}{\qed}
\newcommand{\ra}{\mbox {$\rightarrow$}}
\newtheorem{Theorem}{Theorem}[section]
\newtheorem{Corollary}{Corollary}[section]
\newtheorem{Lemma}{Lemma}[section]
\newtheorem{Example}{Example}[section]
\newenvironment{Proof}{\noindent{\bf Proof:}}{\QED}
\newtheorem{defn}{Definition}
\newtheorem{algo}{Algorithm}
\newtheorem{observation}{Observation}
\newtheorem{problem}{Problem}

%\noBox

\marginparwidth 0pt
\oddsidemargin  0pt
\evensidemargin  0pt
\marginparsep 0pt
\topmargin   0pt
\textwidth   6.5in
\textheight  8.5 in

\newcommand{\cheatbaseline}{1.44}
\renewcommand{\baselinestretch}{\cheatbaseline}

\title{\bf The Michigan Benchmark: Towards XML Query Performance Diagnostics}

\eat{
\author{
{\bf H. V. Jagadish} \\
Univ of Michigan\\
jag@eecs.umich.edu
\and
{\bf Jignesh M. Patel}\\
Univ of Michigan\\
jignesh@eecs.umich.edu
\and
{\bf Kanda Runapongsa}\\
Univ of Michigan\\
krunapon@eecs.umich.edu
}
}

%\date{}

\begin{document}

\maketitle

This benchmark is designed to help engineers design better XML query processing engines.  The data set and queries of this benchmark are described in the following sections.

\section{Benchmark Data Set}
\input dataSet

\section{Benchmark queries}
\label{sec:queries}
\input queries

\bibliographystyle{plain}
\renewcommand{\baselinestretch}{1.1}
\begin{small}
\bibliography{ref}
%\bibliography{lib/bib/moredb,lib/bib/hier,lib/bib/dbimp,lib/bib/text,koudas}
%\begin{thebibliography}{10}

\eat{
\bibitem{PMA}
A.~Apostolico and Z.~Galil.
\newblock Pattern matching algorithms.
\newblock {\em Oxford University Press}, 1997.

\bibitem{Bertino91}
E.~Bertino.
\newblock An indexing technique for object oriented databases.
\newblock {\em ICDE}, pages 14--22, 1991.

\bibitem{bertino}
E.~Bertino and W.~Kim.
\newblock Indexing techniques for queries on nested objects.
\newblock {\em IEEE Transactions on Knowledge and Data Engineering 1(2)}, pages
  196--214, 1989.


\bibitem{bpsm98:xml}
T.~Bray, J.~Paoli, and C.~M. Sperberg-McQueen.
\newblock Extensible markup language ({XML}) 1.0.
\newblock W3C Recommendation. Available at
  http://www.w3.org/TR/1998/REC-xml-19980210, Feb. 1998.
}
\eat{
\bibitem{Jay3}
M.~Carey, J.~Kierman, J.~Shanmugasundaram, E.~Shekita, and S.~Subramanian.
\newblock XPERANTO: Middleware for publishing object relational data as XML
  documents.
\newblock {\em Proceedings of VLDB}, pages 646--648, 2000.

\bibitem{cd+94:shore}
M.~J. Carey, D.~J. DeWitt, M.~J. Franklin, N.~E. Hall, M.~L. McAuliffe, J.~F.
  Naughton, D.~T. Schuh, M.~H. Solomon, C.~K. Tan, O.~G. Tsatalos, S.~J. White,
  and M.~J. Zwilling.
\newblock Shoring up persistent applications.
\newblock In {\em Proceedings of the ACM SIGMOD Conference on Management of
  Data}, pages 383--394, 1994.

\bibitem{crf00:quilt}
D.~D. Chamberlin, J.~Robie, and D.~Florescu.
\newblock Quilt: An {XML} query language for heterogeneous data sources.
\newblock In {\em WebDB (Informal Proceedings)}, pages 53--62, 2000.

\bibitem{dffls98:xmlql}
A.~Deutsch, M.~Fernandez, D.~Florescu, A.~Levy, and D.~Suciu.
\newblock {XML-QL}: A query language for {XML}.
\newblock Submission to the World Wide Web Consortium 19-August-1998. Available
  from http://www.w3.org/TR/NOTE-xml-ql., 1998.

\bibitem{Dewitt91}
D.~DeWitt, J.~Naughton, and D.~Schneider.
\newblock An evaluation of non equijoin algorithms.
\newblock {\em Proceedings of ACM SIGMOD}, pages 443--452, 1991.

\bibitem{Fernandez}
M.~Fernandez and D.~Suciu.
\newblock SilkRoute: Trading between relations and XML.
\newblock {\em WWW9}, 2000.

\bibitem{Moerkotte}
T.~Fiebig and G.~Moerkotte.
\newblock Evaluating queries on structure with access support relations.
\newblock {\em Proceedings of WebDB}, 2000.

\bibitem{fk99:rdbxml}
D.~Florescu and D.~Kossman.
\newblock Storing and querying {XML} data using an {RDBMS}.
\newblock {\em IEEE Data Engineering Bulletin}, 22(3):27--34, 1999.

\bibitem{GraefeQueryEvaluation}
G.~Graefe.
\newblock Query evaluation techniques for large databases.
\newblock {\em ACM Computing Surveys}, 25(2):73--170, 1993.

\bibitem{jkss98:ds}
G.~Jacobson, B.~Krishnamurthy, D.~Srivastava, and D.~Suciu.
\newblock Focusing search in hierarchical structures with directory sets.
\newblock In {\em Proceedings of the Seventh International Conference on
  Information and Knowledge Management (CIKM)}, Washington, DC, Nov. 1998.

\bibitem{jlmsv99:qnd}
H.~V. Jagadish, L.~V.~S. Lakshmanan, T.~Milo, D.~Srivastava, and D.~Vista.
\newblock Querying network directories.
\newblock In {\em Proceedings of the ACM SIGMOD Conference on Management of
  Data}, Philadelphia, PA, June 1999.

\bibitem{kilger}
C.~Kilger and G.~Moerkotte.
\newblock Indexing multiple sets.
\newblock {\em Proceedings of VLDB}, pages 180--191, Sept. 1994.

\bibitem{koudas97}
N.~Koudas and K.~C. Sevcik.
\newblock Size separation spatial join.
\newblock {\em Proceedings of ACM SIGMOD}, pages 324--335, May 1997.

\bibitem{Lo96}
M.-L. Lo and C.~V. Ravishankar.
\newblock Spatial hash-joins.
\newblock {\em Proceedings of ACM SIGMOD}, pages 247--258, June 1996.

\bibitem{McHugh}
J.~McHugh, S.~Abiteboul, R.~Goldman, D.~Quass, and J.~Widom.
\newblock Lore: A database management systems for semistructured data.
\newblock {\em SIGMOD Record 26(3)}, pages 54--66, 1997.


\bibitem{mw99:xml-queryopt}
J.~McHugh and J.~Widom.
\newblock Query optimization for {XML}.
\newblock In {\em Proceedings of the International Conference on Very Large
  Databases}, pages 315--326, 1999.
}
\eat{
\bibitem{tukwila}
U.~of~Washington.
\newblock The {Tukwila} system.
\newblock Available from http://data.cs.washington.edu/integration/tukwila/.

\bibitem{niagara}
U.~of~Wisconsin.
\newblock The {Niagara} system.
\newblock Available from http://www.cs.wisc.edu/niagara/.

\bibitem{Patel96}
J.~M. Patel and D.~J. DeWitt.
\newblock Partition based spatial-merge join.
\newblock {\em Proceedings of ACM SIGMOD}, pages 259--270, June 1996.

\bibitem{Quass96}
D.~Quass, J.~Widom, R.~Goldman, H.~K, Q.~Luo, J.~McHugh, A.~Rajaraman,
  H.~Rivero, S.~. Abiteboul, J.~Ullman, and J.~Wiener.
\newblock Lore: A lightweight object repository for semistructured data.
\newblock {\em Proceedings of ACM SIGMOD}, page 549, 1996.

\bibitem{rls98:xql}
J.~Robie, J.~Lapp, and D.~Schach.
\newblock {XML} query language ({XQL}).
\newblock Available from http://www.w3.org/TandS/QL/QL98/pp/xql.html.

\bibitem{s:kweelt}
A.~Sahuguet.
\newblock Kweelt.
\newblock Available from http://db.cis.upenn.edu/Kweelt/.

\bibitem{sm83:ir}
G.~Salton and M.~J. McGill.
\newblock {\em Introduction to modern information retrieval}.
\newblock McGraw-Hill, New York, 1983.

\bibitem{ssb+2000:xmlpublish}
J.~Shanmugasundaram, E.~J. Shekita, R.~Barr, M.~J. Carey, B.~G.~Lindsay,
  H.~Pirahesh, and B.~Reinwald.
\newblock Efficiently publishing relational data as {XML} documents.
\newblock In {\em Proceedings of the International Conference on Very Large
  Databases}, 2000.

\bibitem{stz+99:rdbxml}
J.~Shanmugasundaram, K.~Tufte, C.~Zhang, G.~He, D.~J. DeWitt, and J.~F.
  Naughton.
\newblock Relational databases for querying {XML} documents: Limitations and
  opportunities.
\newblock In {\em Proceedings of the International Conference on Very Large
  Databases}, 1999.

\bibitem{Shekita90}
E.~Shekita and M.~Carey.
\newblock A performance evaluation of pointer based joins.
\newblock {\em Proceedings of ACM SIGMOD}, pages 300--311, 1990.

\bibitem{sreenath}
B.~Sreenath and S.~Seshadri.
\newblock The hCtree: An efficient index structure for object oriented
  databases.
\newblock {\em Proceedings of VLDB}, pages 203--213, Sept. 1994.

\bibitem{Stein}
J.~Stein and D.~Maier.
\newblock Associative access support in GemStone.
\newblock {\em In K. Dittrich and A. Buchmann, On Object Oriented Database
  systems, Springer,Verlag}, pages 323--339, 1991.

\bibitem{znd+2001:contain}
C.~Zhang, J.~Naughton, D.~Dewitt, Q.~Luo, and G.~Lohman.
\newblock On supporting containment queries in relational database management
  systems.
\newblock In {\em Proceedings of the ACM SIGMOD Conference on Management of
  Data}, 2001.
}

%\end{thebibliography}
\end{small}
\renewcommand{\baselinestretch}{\cheatbaseline}

\end{document}
